---
title: "cML-LFM Tutorial"
output: html_document
date: "`r Sys.Date()`"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
# Install missing packages
list.of.packages <- c("common", "dplyr", "fda", "mgcv", "pracma", "tidyverse")
new.packages <- list.of.packages[!(list.of.packages %in%
                                     installed.packages()[,"Package"])]
if(length(new.packages)) install.packages(new.packages)
library(common)
library(dplyr)
library(fda)
library(mgcv)
library(pracma)
library(tidyverse)
rm(new.packages, list.of.packages)
```




# Introduction

## Description

A step-by-step implementation of cML-LFM and associated procedures including data 
generating process of the first simulation scenario and evaluation of model 
parameter estimation. Details of simulation design and estimation of model parameters
are described in 'Contrastive Latent Models for Structured Functional Data'.

## Main functions implemented

1. data_gen: Function that generates a pair of data sets composed of preset shared and unique variation patterns.
2. cML-LFM: Function that constructs and fits the proposed contrastive multilevel latent functional model.

## Required files

1. Simulation_dgp.R
2. cML-LFM_Functions.R


# Implementation

## 1. Load the required files

```{r, warning=FALSE}
#setwd("C:/Users/Tim/OneDrive - UCLA IT Services/Proj23_SDSI_paper/CodeReleasing/cML-LFM/")
setwd("/Users/a123456/Library/CloudStorage/OneDrive-UCLAITServices/Proj23_SDSI_paper/CodeReleasing/cML-LFM/")
source("Simulation_dgp.R")
source("cML-LFM_Functions.R")
```


## 2. cML-LFM implementation

```{r}
# Containers to save the estimates of mean functions, error variances, 
# latent components, score variances.
mu_x_save = list()
mu_y_save = list()
s2_save = list()

Psi_1_save = list()
Phi_1_save = list()
Gamma_1_save = list()
Psi_2_save = list()
Phi_2_save = list()
Gamma_2_save = list()

Omega_x_1_save = list()
Omega_y_1_save = list()
Theta_1_save = list()
Lambda_1_save = list()
Omega_x_2_save = list()
Omega_y_2_save = list()
Theta_2_save = list()
Lambda_2_save = list()
```

```{R}
for (ite in 1:5) {
  print(paste0(ite, "th iteration starts here ------------------------------"))
  # Generate the simulation data pairs
  data_grids = data_gen(n_x = 50, n_y = 50, s2_x = .25, s2_y = .25, seed = ite)
  X.df = data_grids[[1]]
  Y.df = data_grids[[2]]
  
  # Fitting the proposed cML-LFM 
  results <- cML_LFM(X.df, Y.df)

  # Save the estimates
  mu_x_save = append(mu_x_save, list(results[[1]]))
  mu_y_save = append(mu_y_save, list(results[[2]]))
  s2_save = append(s2_save, list(c(results[[3]], results[[4]])))

  Psi_1_save = append(Psi_1_save, list(results[[5]]))
  Phi_1_save = append(Phi_1_save, list(results[[6]]))
  Gamma_1_save = append(Gamma_1_save, list(results[[7]]))
  Psi_2_save = append(Psi_2_save, list(results[[8]]))
  Phi_2_save = append(Phi_2_save, list(results[[9]]))
  Gamma_2_save = append(Gamma_2_save, list(results[[10]]))

  Omega_x_1_save = append(Omega_x_1_save, list(results[[11]]))
  Omega_y_1_save = append(Omega_y_1_save, list(results[[12]]))
  Theta_1_save = append(Theta_1_save, list(results[[13]]))
  Lambda_1_save = append(Lambda_1_save, list(results[[14]]))
  Omega_x_2_save = append(Omega_x_2_save, list(results[[15]]))
  Omega_y_2_save = append(Omega_y_2_save, list(results[[16]]))
  Theta_2_save = append(Theta_2_save, list(results[[17]]))
  Lambda_2_save = append(Lambda_2_save, list(results[[18]]))
}
```



## 3. Evaluation

True model parameters
```{r, warning=FALSE}
N = 50
tobs = seq(0, 1, length.out = N)

# Error variances
s2_x = .25
s2_y = .25

# Dimension of latent component spaces
L_1 = 2 # dimension of the level-1 shared space
K_1 = 1 # dimension of the level-1 unique space of the first group of data X
R_1 = 1 # dimension of the level-1 unique space of the second group of data Y
L_2 = 1 # dimension of the level-2 shared space
K_2 = 2 # dimension of the level-2 unique space of the first group of data X
R_2 = 1 # dimension of the level-2 unique space of the second group of data Y


# Latent components
## Level-1
Psi_1 = eval.basis(tobs, suppressWarnings(
  create.fourier.basis(nbasis = 5, dropind = c(1,3,5)))) 
Phi_1 = eval.basis(tobs, suppressWarnings(
  create.fourier.basis(nbasis = 3, dropind = c(1,2))))
Gamma_1 = eval.basis(tobs, suppressWarnings(
  create.fourier.basis(nbasis = 5, dropind = c(1:4))))

## Level-2
Psi_2 = matrix(sapply(tobs, FUN = function(x) 1), ncol = 1)
Phi_2 = cbind(matrix(sapply(tobs, FUN = function(x) (sqrt(3)*(2*x-1))), 
                     ncol = 1),
              matrix(sapply(tobs, FUN = function(x) (sqrt(5)*(6*x^2-6*x+1))), 
                     ncol = 1)
              )
Gamma_2 = matrix(sapply(tobs, FUN = function(x) (sqrt(7)*
                                                   (20*x^3-30*x^2+12*x-1))), 
                 ncol = 1)

# Score variances
## Level-1
Omega_x_1 = diag(c(1,.5))
Omega_y_1 = diag(c(1,.5))
Theta_1 = matrix(.25)
Lambda_1 = matrix(.25)

## Level-2
Omega_x_2 = matrix(1)
Omega_y_2 = matrix(1)
Theta_2 = diag(c(.5,.25))
Lambda_2 = matrix(.5)
```



MSE of error variance $\sigma^2_X$, $\sigma^2_Y$:
```{r}
MSE_s2 = matrix(data = NA, nrow = length(s2_save), ncol = 2)
for (i in 1:length(s2_save)) {
  MSE_s2[i,1] = (s2_x - s2_save[[i]][1])^2 / s2_x^2
  MSE_s2[i,2] = (s2_y - s2_save[[i]][2])^2 / s2_y^2
}
print(paste0("The mean MSE for the error variances {", 
             symbol("sigma") %p% supsc("2"), "X, ", 
             symbol("sigma") %p% supsc("2"), "Y} are:"))
print(round(colMeans(MSE_s2, na.rm = TRUE), digits = 3))
```


Calculate MSDE for latent component estimates and MSE for score variance estimates, and plot true and 
estimated trajectories with (5th, 50th, 95th) MSDE values.

*Level-1*

$\psi_\ell^{(1)}(t)$
```{r}
MSDE_Psi = matrix(data = NA, nrow = length(Psi_1_save), ncol = L_1)
Psi_1_est = matrix(data = NA, nrow = length(Psi_1_save), ncol = N)
Psi_2_est = matrix(data = NA, nrow = length(Psi_1_save), ncol = N)
Omega_x_est = matrix(data = NA, nrow = length(Omega_x_1_save), ncol = L_1)
Omega_y_est = matrix(data = NA, nrow = length(Omega_y_1_save), ncol = L_1)
for (i in 1:length(Psi_1_save)) {
  Psi_est = Psi_1_save[[i]]
  if(NCOL(Psi_est) >= L_1){ # skipped if estimated shared components are less than the true ones.
    min.ind = c() # used to save the index of estimated component corresponding to psi^{(1)}_l(t)
    for(j.true in 1:L_1){
      MSDE = c()
      for (j in 1:NCOL(Psi_est)) {
        MSDE = append(MSDE, min(
          trapz(tobs, (Psi_1[,j.true] - Psi_est[,j])^2) / 
            trapz(tobs, Psi_1[,j.true]^2),
          trapz(tobs, (Psi_1[,j.true] + Psi_est[,j])^2) / 
            trapz(tobs, Psi_1[,j.true]^2)))
      }
      min.ind = append(min.ind, which.min(MSDE))  
      MSDE_Psi[i, j.true] = min(MSDE)
      Omega_x_est[i,j.true] = Omega_x_1_save[[i]][which.min(MSDE),which.min(MSDE)]
      Omega_y_est[i,j.true] = Omega_y_1_save[[i]][which.min(MSDE),which.min(MSDE)]
      if(j.true == 1){Psi_1_est[i,] = Psi_est[,which.min(MSDE)]}
      if(j.true == 2){Psi_2_est[i,] = Psi_est[,which.min(MSDE)]}
    }
  }
}

print(paste0("The mean MSDE for the shared components {", symbol("psi"), 
             "(t)} ", "are:"))
print(round(colMeans(MSDE_Psi, na.rm = TRUE), digits = 3))


# Plots

## Choose the estimate with (5th, 50th, 95th) MSDE with the true latent component
quantile_mat = matrix(nrow = L_1, ncol = 3) # a matrix saves the (5th, 50th, 95th) MSDE values.
for(j.true in 1:L_1){
  quantile_mat[j.true,] = sort(na.omit(MSDE_Psi)[,j.true])[ceiling(
    nrow(na.omit(MSDE_Psi)) * c(.05, .50, .95))] 
}

ind_mat = matrix(nrow = L_1, ncol = 3) # a matrix contains the indices corresponding to the selected MSDE
for(j.true in 1:L_1){
  for (j in 1:3) {
    ind_mat[j.true,j] = which.min(abs(MSDE_Psi[,j.true] - quantile_mat[j.true,j]))
  }
}


## plot for psi^{(1)}_1(t)
psi_1_5th = Psi_1_est[ind_mat[1,1],] * sign(Psi_1_est[ind_mat[1,1],11])
psi_1_50th = Psi_1_est[ind_mat[1,2],] * sign(Psi_1_est[ind_mat[1,2],11])
psi_1_95th = Psi_1_est[ind_mat[1,3],] * sign(Psi_1_est[ind_mat[1,3],11])
matplot(tobs, cbind(Psi_1[,1], psi_1_5th, psi_1_50th, psi_1_95th), 
        type = "l", lwd = 3, xlab = "Time", ylab = "",  cex.lab = 1.9, 
        cex.axis = 1.5, col=c("black", "red","black", "blue"), lty = c(1,2,3,4), 
        mgp = c(2.5,1,0), xaxs = "i")
title(main = list("", cex = 2.5), adj = 0, line = 1.5)
mtext(expression(hat(psi)[1]^{(1)}*(t)), side = 2, line = 1.9, cex = 1.7)
legend("bottomleft", c("true", "5th", "50th", "95th"), lty = c(1,2,3,4), 
       col=c("black", "red", "black", "blue"), lwd = 2, cex=1)


## plot for psi^{(1)}_2(t)
if(L_1 > 1){
  psi_2_5th = Psi_2_est[ind_mat[2,1],] * sign(Psi_2_est[ind_mat[2,1],6])
  psi_2_50th = Psi_2_est[ind_mat[2,2],] * sign(Psi_2_est[ind_mat[2,2],6])
  psi_2_95th = Psi_2_est[ind_mat[2,3],] * sign(Psi_2_est[ind_mat[2,3],6])
  matplot(tobs, cbind(Psi_1[,2], psi_2_5th, psi_2_50th, psi_2_95th), 
          type = "l", lwd = 3, xlab = "Time", ylab = "",  cex.lab = 1.9, 
          cex.axis = 1.5, col=c("black", "red","black", "blue"), 
          lty = c(1,2,3,4), mgp = c(2.5,1,0), xaxs = "i")
  title(main = list("", cex = 2.5), adj = 0, line = 1.5)
  mtext(expression(hat(psi)[2]^{(1)}*(t)), side=2, line=1.9, cex=1.9)
}
```



$\phi_k^{(1)}(t)$
```{r}
MSDE_Phi = matrix(data = NA, nrow = length(Phi_1_save), ncol = K_1)
Phi_1_est = matrix(data = NA, nrow = length(Phi_1_save), ncol = N)
Theta_est = matrix(data = NA, nrow = length(Theta_1_save), ncol = K_1)
for (i in 1:length(Phi_1_save)) {
  Phi_est = Phi_1_save[[i]]
  if(NCOL(Phi_est) >= K_1){
    if(NCOL(Phi_est) == 1){ # K_1 = K_est = 1.
      MSDE_Phi[i,1] = min(
        trapz(tobs, (Phi_1 - Phi_est)^2) / trapz(tobs, Phi_1^2), 
        trapz(tobs, (Phi_1 + Phi_est)^2) / trapz(tobs, Phi_1^2))
      Phi_1_est[i,] = Phi_est
      Theta_est[i,1] = Theta_1_save[[i]]
    }else{ # K_est > K_1 = 1
      # When surplus unique component estimates are returned, the one having the 
      # lowest MSDE with the true component is referred to as its correspondence.
      for (j in 1:NCOL(Phi_est)) {
        MSDE = min(trapz(tobs, (Phi_1 - Phi_est[,j])^2) / 
                     trapz(tobs, Phi_1^2), 
                 trapz(tobs, (Phi_1 + Phi_est[,j])^2) / 
                   trapz(tobs, Phi_1^2))
        if(j > 1){
          if(MSDE < MSDE_Phi[i, 1]){
            MSDE_Phi[i, 1] = MSDE
            Phi_1_est[i,] = Phi_est[,j]
            Theta_est[i,1] = Theta_1_save[[i]][j,j]
          }
        }else{
          MSDE_Phi[i, 1] = MSDE
          Phi_1_est[i,] = Phi_est[,j]
          Theta_est[i,1] = Theta_1_save[[i]][j,j]
        }
      }
    }
  }
}


print(paste0("The mean MSDE for the unique components {", symbol("phi"), 
             "(t)} ", "are:"))
print(round(colMeans(MSDE_Phi, na.rm = TRUE), digits = 3))



# Plots

quantile_mat = matrix(nrow = K_1, ncol = 3)
for(j.true in 1:K_1){
  quantile_mat[j.true,] = sort(na.omit(MSDE_Phi)[,j.true])[ceiling(nrow(na.omit(MSDE_Phi)) * 
                                                       c(.05, .50, .95))] 
}
ind_mat = matrix(nrow = K_1, ncol = 3) 
for(j.true in 1:K_1){
  for (j in 1:3) {
    ind_mat[j.true,j] = which.min(abs(MSDE_Phi[,j.true] - quantile_mat[j.true,j]))
  }
}


## plot for phi^{(1)}_1(t)
phi_1_5th = Phi_1_est[ind_mat[1,1],] * sign(Phi_1_est[ind_mat[1,1],1])
phi_1_50th = Phi_1_est[ind_mat[1,2],] * sign(Phi_1_est[ind_mat[1,2],1])
phi_1_95th = Phi_1_est[ind_mat[1,3],] * sign(Phi_1_est[ind_mat[1,3],1])
matplot(tobs, cbind(Phi_1[,1], phi_1_5th, phi_1_50th, phi_1_95th), 
        type = "l", lwd = 3, xlab = "Time", ylab = "", cex.lab = 1.9, 
        cex.axis = 1.5, col=c("black", "red","black", "blue"), lty = c(1,2,3,4), 
        mgp = c(2.5,1,0), , xaxs = "i")
title(main = list("", cex = 2.5), adj = 0, line = 1.5)
mtext(expression(hat(phi)[1]^{(1)}*(t)), side=2, line=1.9, cex=1.4)
```



$\gamma_r^{(1)}(t)$
```{r}
MSDE_Gamma = matrix(data = NA, nrow = length(Gamma_1_save), ncol = R_1)
Gamma_1_est = matrix(data = NA, nrow = length(Gamma_1_save), ncol = N)
Lambda_est = matrix(data = NA, nrow = length(Lambda_1_save), ncol = R_1)
for (i in 1:length(Gamma_1_save)) {
  Gamma_est = Gamma_1_save[[i]]
  if(NCOL(Gamma_est) >= R_1){
    if(NCOL(Gamma_est) == 1){ # R_1 = R_est = 1.
      MSDE_Gamma[i,1] = min(
        trapz(tobs, (Gamma_1 - Gamma_est)^2) / trapz(tobs, Gamma_1^2), 
        trapz(tobs, (Gamma_1 + Gamma_est)^2) / trapz(tobs, Gamma_1^2))
      Gamma_1_est[i,] = Gamma_est
      Lambda_est[i,1] = Lambda_1_save[[i]]
    }else if(R_1 == 1){ # R_est > R_1 = 1
      for (j in 1:NCOL(Gamma_est)) {
        MSDE = min(trapz(tobs, (Gamma_1 - Gamma_est[,j])^2) / 
                     trapz(tobs, Gamma_1^2), 
                 trapz(tobs, (Gamma_1 + Gamma_est[,j])^2) / 
                   trapz(tobs, Gamma_1^2))
        if(j > 1){
          if(MSDE < MSDE_Gamma[i, 1]){
            MSDE_Gamma[i,1] = MSDE
            Gamma_1_est[i,] = Gamma_est[,j]
            Lambda_est[i,1] = Lambda_1_save[[i]][j,j]
          }
        }else{
          MSDE_Gamma[i,1] = MSDE
          Gamma_1_est[i,] = Gamma_est[,j]
          Lambda_est[i,1] = Lambda_1_save[[i]][j,j]
        }
      }
    }
  }
}

print(paste0("The mean MSDE for the unique components {", symbol("gamma"), 
             "(t)} ", "are:"))
print(round(colMeans(MSDE_Gamma, na.rm = TRUE), digits = 3))


# Plots

quantile_mat = matrix(nrow = R_1, ncol = 3) 
for(j.true in 1:R_1){
  quantile_mat[j.true,] = sort(na.omit(MSDE_Gamma)[,j.true])[ceiling(nrow(na.omit(MSDE_Gamma)) * 
                                                       c(.05, .50, .95))] 
}
ind_mat = matrix(nrow = R_1, ncol = 3) 
for(j.true in 1:R_1){
  for (j in 1:3) {
    ind_mat[j.true,j] = which.min(abs(MSDE_Gamma[,j.true] - quantile_mat[j.true,j]))
  }
}

## plot for gamma^{(1)}_1(t)
gamma_1_5th = Gamma_1_est[ind_mat[1,1],] * sign(Gamma_1_est[ind_mat[1,1],1])
gamma_1_50th = Gamma_1_est[ind_mat[1,2],] * sign(Gamma_1_est[ind_mat[1,2],1])
gamma_1_95th = Gamma_1_est[ind_mat[1,3],] * sign(Gamma_1_est[ind_mat[1,3],1])
matplot(tobs, cbind(Gamma_1[,1], gamma_1_5th, gamma_1_50th, gamma_1_95th), 
        type = "l", lwd = 3, xlab = "Time", ylab = "", cex.lab = 1.9, 
        cex.axis = 1.5, col=c("black", "red","black", "blue"), lty = c(1,2,3,4),
        mgp = c(2.5,1,0), xaxs = "i")
title(main = list("", cex = 2.5), adj = 0, line = 1.5)
mtext(expression(hat(gamma)[1]^{(1)}*(t)), side=2, line=1.9, cex=1.5)
```


Calculate MSE of score variances

```{r}
scoreVar_MSE = matrix(data = NA, nrow = length(Omega_x_1_save), ncol = (2*L_1 + K_1 + R_1))
for (i in 1:length(Omega_x_1_save)) {
    scoreVar_MSE[i,1:L_1] = (diag(Omega_x_1) - Omega_x_est[i,])^2 / (diag(Omega_x_1))^2
    scoreVar_MSE[i,(L_1+1):(L_1+L_1)] = (diag(Omega_y_1) - Omega_y_est[i,])^2 / (diag(Omega_y_1))^2
    scoreVar_MSE[i,(2*L_1+1):(2*L_1+K_1)] = (diag(Theta_1) - Theta_est[i,])^2 / (diag(Theta_1))^2
    scoreVar_MSE[i,(2*L_1+K_1+1):(2*L_1+K_1+R_1)] = (diag(Lambda_1) - Lambda_est[i,])^2 / (diag(Lambda_1))^2
}

print("Mean MSE for the following score variances are:")
print(paste(symbol("omega") %p% supsc("x"), ":", 
            round(colMeans(as.matrix(scoreVar_MSE[,1:L_1]), na.rm = TRUE), digits = 3)))
print(paste(symbol("omega") %p% supsc("y"), ":", 
            round(colMeans(as.matrix(scoreVar_MSE[,(L_1+1):(L_1+L_1)]), na.rm = TRUE), digits = 3)))
print(paste(symbol("theta"), ":", 
            round(colMeans(as.matrix(scoreVar_MSE[,(2*L_1+1):(2*L_1+K_1)]), na.rm = TRUE), digits = 3)))
print(paste(symbol("lambda"), ":", 
            round(colMeans(as.matrix(scoreVar_MSE[,(2*L_1+K_1+1):(2*L_1+K_1+R_1)]), na.rm = TRUE), digits = 3)))
```



*Level-2*

$\psi_\ell^{(2)}(t)$
```{r}
MSDE_Psi = matrix(data = NA, nrow = length(Psi_2_save), ncol = L_2)
Psi_1_est = matrix(data = NA, nrow = length(Psi_2_save), ncol = N)
Omega_x_est = matrix(data = NA, nrow = length(Omega_x_2_save), ncol = L_2)
Omega_y_est = matrix(data = NA, nrow = length(Omega_y_2_save), ncol = L_2)
for (i in 1:length(Psi_2_save)) {
  Psi_est = Psi_2_save[[i]]
  if(NCOL(Psi_est) >= L_2){ # skipped if no shared components are estimated.
    if(NCOL(Psi_est) == L_2){ 
      MSDE_Psi[i,1] = min(
        trapz(tobs, (Psi_2 - Psi_est)^2)/trapz(tobs, Psi_2^2), 
        trapz(tobs, (Psi_2 + Psi_est)^2)/trapz(tobs, Psi_2^2))
      Psi_1_est[i,] = Psi_est
      Omega_x_est[i,1] = Omega_x_2_save[[i]]
      Omega_y_est[i,1] = Omega_y_2_save[[i]]
    }else{ # L_est > L_2 = 1
      for (j in 1:NCOL(Psi_est)) {
        MSDE = min(trapz(tobs, (Psi_2 - Psi_est[,j])^2) / trapz(tobs, Psi_2^2), 
                 trapz(tobs, (Psi_2 + Psi_est[,j])^2) / trapz(tobs, Psi_2^2))
        if(j > 1){
          if(MSDE < MSDE_Psi[i, 1]){
            MSDE_Psi[i, 1] = MSDE
            Psi_1_est[i,] = Psi_est[,j]
            Omega_x_est[i,1] = Omega_x_2_save[[i]][j,j]
            Omega_y_est[i,1] = Omega_y_2_save[[i]][j,j]
          }
        }else{
          MSDE_Psi[i, 1] = MSDE
          Psi_1_est[i,] = Psi_est[,j]
          Omega_x_est[i,1] = Omega_x_2_save[[i]][j,j]
          Omega_y_est[i,1] = Omega_y_2_save[[i]][j,j]
        }
      }
    }
  }
}

print(paste0("The mean MSDE for the shared components {", symbol("psi"), 
             "(t)} ", "are:"))
print(round(colMeans(MSDE_Psi, na.rm = TRUE), digits = 3))

# Plots

## Choose the estimate with (5th, 50th, 95th) MSDE with the true latent component
quantile_mat = matrix(nrow = L_2, ncol = 3) # a matrix saves the (5th, 50th, 95th) MSDE values.
for(j.true in 1:L_2){
  quantile_mat[j.true,] = sort(na.omit(MSDE_Psi)[,j.true])[ceiling(nrow(na.omit(MSDE_Psi)) * 
                                                       c(.05, .50, .95))] 
}


ind_mat = matrix(nrow = L_2, ncol = 3) # a matrix contains the indices corresponding to the selected MSDE
for(j.true in 1:L_2){
  for (j in 1:3) {
    ind_mat[j.true,j] = which.min(abs(MSDE_Psi[,j.true] - quantile_mat[j.true,j]))
  }
}

## plot for psi^{(2)}_1(t)
psi_1_5th = Psi_1_est[ind_mat[1,1],] * sign(Psi_1_est[ind_mat[1,1],1])
psi_1_50th = Psi_1_est[ind_mat[1,2],] * sign(Psi_1_est[ind_mat[1,2],1])
psi_1_95th = Psi_1_est[ind_mat[1,3],] * sign(Psi_1_est[ind_mat[1,3],1])

par(mar = c(5.1, 5.2, 4.1, 2.1))
matplot(tobs, cbind(Psi_2[,1], psi_1_5th, psi_1_50th, psi_1_95th), 
        type = "l", lwd = 3, xlab = "Time", ylab = "",  cex.lab = 1.9, 
        cex.axis = 1.5, col=c("black", "red","black", "blue"), lty = c(1,2,3,4), 
        mgp = c(2.5,1,0), xaxs = "i", ylim = c(-.25, 1.75))
title(main = list("", cex = 2.5), adj = 0, line = 1.5)
mtext(expression(hat(psi)[1]^{(2)}*(t)), side=2, line=2.1, cex=2)
legend("bottomleft", c("true", "5th", "50th", "95th"), lty = c(1,2,3,4), 
       col=c("black", "red", "black", "blue"), lwd = 2, cex=1.2)
```


$\phi_k^{(2)}(t)$
```{r}
MSDE_Phi = matrix(data = NA, nrow = length(Phi_2_save), ncol = K_2)
Phi_1_est = matrix(data = NA, nrow = length(Phi_2_save), ncol = N)
Phi_2_est = matrix(data = NA, nrow = length(Phi_2_save), ncol = N)
Theta_est = matrix(data = NA, nrow = length(Theta_2_save), ncol = K_2)
for (i in 1:length(Phi_2_save)) {
  Phi_est = Phi_2_save[[i]]
  if(NCOL(Phi_est) >= K_2){
    min.ind = c()
    for(j.true in 1:K_2){
      MSDE = c()
      for (j in 1:NCOL(Phi_est)) {
        MSDE = append(MSDE, min(
          trapz(tobs, (Phi_2[,j.true] - Phi_est[,j])^2) / 
            trapz(tobs, Phi_2[,j.true]^2),
          trapz(tobs, (Phi_2[,j.true] + Phi_est[,j])^2) / 
            trapz(tobs, Phi_2[,j.true]^2)))
      }
      min.ind = append(min.ind, which.min(MSDE))  
      MSDE_Phi[i, j.true] = min(MSDE)
      Theta_est[i,j.true] = Theta_2_save[[i]][which.min(MSDE),which.min(MSDE)]
      if(j.true == 1){Phi_1_est[i,] = Phi_est[,which.min(MSDE)]}
      if(j.true == 2){Phi_2_est[i,] = Phi_est[,which.min(MSDE)]}
    }
  }
}

print(paste0("The mean MSDE for the unique components {", symbol("phi"), 
             "(t)} ", "are:"))
print(round(colMeans(MSDE_Phi, na.rm = TRUE), digits = 3))


# Plots

## Choose the estimate with (5th, 50th, 95th) MSDE with the true latent component
quantile_mat = matrix(nrow = K_2, ncol = 3) # a matrix saves the (5th, 50th, 95th) MSDE values.
for(j.true in 1:K_2){
  quantile_mat[j.true,] = sort(na.omit(MSDE_Phi)[,j.true])[ceiling(
    nrow(na.omit(MSDE_Phi)) * c(.05, .50, .95))] 
}

ind_mat = matrix(nrow = K_2, ncol = 3) # a matrix contains the indices corresponding to the selected MSDE
for(j.true in 1:K_2){
  for (j in 1:3) {
    ind_mat[j.true,j] = which.min(abs(MSDE_Phi[,j.true] - quantile_mat[j.true,j]))
  }
}


## plot for phi^{(2)}_1(t)
phi_1_5th = Phi_1_est[ind_mat[1,1],] * sign(Phi_1_est[ind_mat[1,1],50])
phi_1_50th = Phi_1_est[ind_mat[1,2],] * sign(Phi_1_est[ind_mat[1,2],50])
phi_1_95th = Phi_1_est[ind_mat[1,3],] * sign(Phi_1_est[ind_mat[1,3],50])

par(mar = c(5.1, 5.2, 4.1, 2.1))
matplot(tobs, cbind(Phi_2[,1], phi_1_5th, phi_1_50th, phi_1_95th), 
        type = "l", lwd = 3, xlab = "Time", ylab = "", cex.lab = 1.9, 
        cex.axis = 1.5, col=c("black", "red","black", "blue"), lty = c(1,2,3,4), 
        mgp = c(2.5,1,0), , xaxs = "i")
title(main = list("", cex = 2.5), adj = 0, line = 1.5)
mtext(expression(hat(phi)[1]^{(2)}*(t)), side=2, line=2.1, cex=2)


## plot for phi^{(2)}_2(t)
phi_2_5th = Phi_2_est[ind_mat[2,1],] * sign(Phi_2_est[ind_mat[2,1],1])
phi_2_50th = Phi_2_est[ind_mat[2,2],] * sign(Phi_2_est[ind_mat[2,2],1])
phi_2_95th = Phi_2_est[ind_mat[2,3],] * sign(Phi_2_est[ind_mat[2,3],1])

par(mar = c(5.1, 5.2, 4.1, 2.1))
matplot(tobs, cbind(Phi_2[,2], phi_2_5th, phi_2_50th, phi_2_95th),
        type = "l", lwd = 3, xlab = "Time", ylab = "",  cex.lab = 1.9,
        cex.axis = 1.5, col=c("black", "red","black", "blue"),
        lty = c(1,2,3,4), mgp = c(2.5,1,0), , xaxs = "i")
title(main = list("", cex = 2.5), adj = 0, line = 1.5)
mtext(expression(hat(phi)[2]^{(2)}*(t)), side=2, line=2.1, cex=2)
```


$\gamma_r^{(2)}(t)$
```{r}
MSDE_Gamma = matrix(data = NA, nrow = length(Gamma_2_save), ncol = R_2)
Gamma_1_est = matrix(data = NA, nrow = length(Gamma_2_save), ncol = N)
Lambda_est = matrix(data = NA, nrow = length(Lambda_2_save), ncol = R_2)
for (i in 1:length(Gamma_2_save)) {
  Gamma_est = Gamma_2_save[[i]]
  if(NCOL(Gamma_est) >= R_2){
    if(NCOL(Gamma_est) == R_2){
      MSDE_Gamma[i,1] = min(
        trapz(tobs, (Gamma_2 - Gamma_est)^2) / trapz(tobs, Gamma_2^2), 
        trapz(tobs, (Gamma_2 + Gamma_est)^2) / trapz(tobs, Gamma_2^2))
      Gamma_1_est[i,] = Gamma_est
      Lambda_est[i,1] = Lambda_2_save[[i]]
    }else{ # R_est > R_2 = 1
      # When surplus unique component estimates are returned, the one having the 
      # lowest MSDE with the true component is referred to as its correspondence.
      for (j in 1:NCOL(Gamma_est)) {
        MSDE = min(trapz(tobs, (Gamma_2 - Gamma_est[,j])^2) / 
                     trapz(tobs, Gamma_2^2), 
                 trapz(tobs, (Gamma_2 + Gamma_est[,j])^2) / 
                   trapz(tobs, Gamma_2^2))
        if(j > 1){
          if(MSDE < MSDE_Gamma[i, 1]){
            MSDE_Gamma[i,1] = MSDE
            Gamma_1_est[i,] = Gamma_est[,j]
            Lambda_est[i,1] = Lambda_2_save[[i]][j,j]
          }
        }else{
          MSDE_Gamma[i,1] = MSDE
          Gamma_1_est[i,] = Gamma_est[,j]
          Lambda_est[i,1] = Lambda_2_save[[i]][j,j]
        }
      }
    }
  }
}

print(paste0("The mean MSDE for the unique components {", symbol("gamma"), 
             "(t)} ", "are:"))
print(round(colMeans(MSDE_Gamma, na.rm = TRUE), digits = 3))


# Plots

## Choose the estimate with (5th, 50th, 95th) MSDE with the true latent component
quantile_mat = matrix(nrow = R_2, ncol = 3) # a matrix saves the (5th, 50th, 95th) MSDE values.
for(j.true in 1:R_2){
  quantile_mat[j.true,] = sort(na.omit(MSDE_Gamma)[,j.true])[ceiling(
    nrow(na.omit(MSDE_Gamma)) * c(.05, .50, .95))] 
}
ind_mat = matrix(nrow = R_2, ncol = 3) # a matrix contains the indices corresponding to the selected MSDE
for(j.true in 1:R_2){
  for (j in 1:3) {
    ind_mat[j.true,j] = which.min(abs(MSDE_Gamma[,j.true] - quantile_mat[j.true,j]))
  }
}

## plot for gamma^{(2)}_1(t)
gamma_1_5th = Gamma_1_est[ind_mat[1,1],] * sign(Gamma_1_est[ind_mat[1,1],50])
gamma_1_50th = Gamma_1_est[ind_mat[1,2],] * sign(Gamma_1_est[ind_mat[1,2],50])
gamma_1_95th = Gamma_1_est[ind_mat[1,3],] * sign(Gamma_1_est[ind_mat[1,3],50])

par(mar = c(5.1, 5.2, 4.1, 2.1))
matplot(tobs, cbind(Gamma_2[,1], gamma_1_5th, gamma_1_50th, gamma_1_95th), 
        type = "l", lwd = 3, xlab = "Time", ylab = "", cex.lab = 1.9, 
        cex.axis = 1.5, col=c("black", "red","black", "blue"), lty = c(1,2,3,4),
        mgp = c(2.5,1,0), xaxs = "i")
title(main = list("", cex = 2.5), adj = 0, line = 1.5)
mtext(expression(hat(gamma)[1]^{(2)}*(t)), side=2, line=2.1, cex=2)
```


Calculate MSE of score variances

```{r}
scoreVar_MSE = matrix(data = NA, nrow = length(Omega_x_2_save), ncol = (2*L_2 + K_2 + R_2))
for (i in 1:length(Omega_x_2_save)) {
    scoreVar_MSE[i,1:L_2] = (diag(Omega_x_2) - Omega_x_est[i,])^2 / (diag(Omega_x_2))^2
    scoreVar_MSE[i,(L_2+1):(L_2+L_2)] = (diag(Omega_y_2) - Omega_y_est[i,])^2 / (diag(Omega_y_2))^2
    scoreVar_MSE[i,(2*L_2+1):(2*L_2+K_2)] = (diag(Theta_2) - Theta_est[i,])^2 / (diag(Theta_2))^2
    scoreVar_MSE[i,(2*L_2+K_2+1):(2*L_2+K_2+R_2)] = (diag(Lambda_2) - Lambda_est[i,])^2 / (diag(Lambda_2))^2
}

print("Mean MSE for the following score variances are:")
print(paste(symbol("omega") %p% supsc("x"), ":", 
            round(colMeans(as.matrix(scoreVar_MSE[,1:L_2]), na.rm = TRUE), digits = 3)))
print(paste(symbol("omega") %p% supsc("y"), ":", 
            round(colMeans(as.matrix(scoreVar_MSE[,(L_2+1):(L_2+L_2)]), na.rm = TRUE),digits = 3)))
print(paste(symbol("theta"), ":", 
            round(colMeans(as.matrix(scoreVar_MSE[,(2*L_2+1):(2*L_2+K_2)]), na.rm = TRUE),digits = 3)))
print(paste(symbol("lambda"), ":", 
            round(colMeans(as.matrix(scoreVar_MSE[,(2*L_2+K_2+1):(2*L_2+K_2+R_2)]), na.rm = TRUE),digits = 3)))
```







