---
title: "cMV-LFM Tutorial"
output: html_document
date: "`r Sys.Date()`"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
# Install missing packages
list.of.packages <- c("common", "dplyr", "fda", "Matrix", "mfaces", "mgcv", 
                      "pracma", "tidyverse")
new.packages <- list.of.packages[!(list.of.packages %in%
                                     installed.packages()[,"Package"])]
if(length(new.packages)) install.packages(new.packages)
library(common)
library(dplyr)
library(fda)
library(Matrix)
library(mfaces)
library(mgcv)
library(pracma)
library(tidyverse)
rm(new.packages, list.of.packages)
```




# Introduction

## Description

A step-by-step implementation of cMV-LFM and associated procedures including data 
generating process of the first simulation scenario and evaluation of model 
parameter estimation. Details of simulation design and estimation of model parameters
are described in 'Contrastive Latent Models for Structured Functional Data'.

## Main functions implemented

1. data_gen: Function that generates a pair of data sets composed of preset shared and unique variation patterns.
2. cMV-LFM: Function that constructs and fits the proposed contrastive multivariate latent functional model.

## Required files

1. Simulation_dgp.R
2. cMV-LFM_Functions.R


# Implementation

## 1. Load the required files

```{r,warning=FALSE}
setwd("C:/Users/Tim/OneDrive - UCLA IT Services/SDSI_paper/CodeReleasing/cMV-LFM/")
source("Simulation_dgp.R")
source("cMV-LFM_Functions.R")
```



## 2. cMV-LFM implementation

```{r}
# Containers to save the estimates of mean functions, error variances, scaling parameters,
# latent components and score variances.
mu_x_save = list()
mu_y_save = list()
s2_x_save = list()
s2_y_save = list()
beta.X_save = list()
beta.Y_save = list()

Psi_save = list()
Phi_save = list()
Gamma_save = list()
tau_x_save = list()
tau_y_save = list()

Omega_x_save = list()
Omega_y_save = list()
Theta_save = list()
Lambda_save = list()
varsigma_x_save = list()
varsigma_y_save = list()
```

```{R}
for (ite in 1:5 ) {
  print(paste0(ite, "th iteration starts here ------------------------------"))
  # Generate the simulation data pairs
  data_grids = data_gen(n_x = 50, n_y = 50, s2_x = .25, s2_y = .25, seed = ite)
  X.ls = data_grids[[1]]
  Y.ls = data_grids[[2]]
  
  # Fitting the proposed cMV-LFM 
  results <- cMV_LFM(X.ls, Y.ls)
    
  # Save the estimates
  mu_x_save = append(mu_x_save, list(results[[1]]))
  mu_y_save = append(mu_y_save, list(results[[2]]))
  s2_x_save = append(s2_x_save, list(results[[3]]))
  s2_y_save = append(s2_y_save, list(results[[4]]))
  beta.X_save = append(beta.X_save, list(results[[5]]))
  beta.Y_save = append(beta.Y_save, list(results[[6]]))

  Psi_save = append(Psi_save, list(results[[7]]))
  Phi_save = append(Phi_save, list(results[[8]]))
  Gamma_save = append(Gamma_save, list(results[[9]]))
  tau_x_save = append(tau_x_save, list(results[[10]]))
  tau_y_save = append(tau_y_save, list(results[[11]]))

  Omega_x_save = append(Omega_x_save, list(results[[12]]))
  Omega_y_save = append(Omega_y_save, list(results[[13]]))
  Theta_save = append(Theta_save, list(results[[14]]))
  Lambda_save = append(Lambda_save, list(results[[15]]))
  varsigma_x_save = append(varsigma_x_save, list(results[[16]]))
  varsigma_y_save = append(varsigma_y_save, list(results[[17]]))
}
```


## 3. Evaluation

True model parameters
```{r, warning=FALSE}
N = 50
tobs = seq(0, 1, length.out = N)
J = 12

# Error variances
s2_x = rep(.25, J)
s2_y = rep(.25, J)

# Scaling parameters
beta.X <- c(1, 0.88, 0.84, 0.75, 0.71, 0.93, 0.81, 1.11, 0.80, 0.89, 0.90, 0.73)
beta.Y <- c(1, 0.59, 0.63, 0.54, 0.35, 0.83, 0.73, 0.98, 0.78, 0.92, 0.69, 0.71)

# Dimension of latent component spaces
L = 1 # dimension of the subject-level shared space
K = 1 # dimension of the subject-level unique space of the first group of data X
R = 1 # dimension of the subject-level unique space of the second group of data Y
V_x = 2 # dimension of the outcome-level latent space of X
V_y = 3 # dimension of the outcome-level latent space of Y

# Latent components
## Subject-level
Psi = eval.basis(tobs, suppressWarnings(create.fourier.basis(
  nbasis = 3, dropind = c(1,3)))) 
Phi = eval.basis(tobs, suppressWarnings(create.fourier.basis(
  nbasis = 3, dropind = c(1,2))))
Gamma = eval.basis(tobs, suppressWarnings(create.fourier.basis(
  nbasis = 5, dropind = c(1:3,5))))

## Outcome-level
tau_x = cbind(
  matrix(sapply(tobs, FUN = function(x) 1), ncol = 1),
  matrix(sapply(tobs, FUN = function(x) (sqrt(3)*(2*x-1))), ncol = 1)
) 
tau_y = cbind(
  matrix(sapply(tobs, FUN = function(x) 1), ncol = 1),
  matrix(sapply(tobs, FUN = function(x) (sqrt(5)*(6*x^2-6*x+1))), ncol = 1),
  matrix(sapply(tobs, FUN = function(x) (sqrt(7)*(20*x^3-30*x^2+12*x-1))), ncol = 1)
) 

# Score variances
## Subject-level
Omega_x = matrix(1)
Omega_y = matrix(1)
Theta = matrix(.5)
Lambda = matrix(.5)

## Outcome-level
varsigma_x = diag(c(1, .5))
varsigma_y = diag(c(1, .5, .25))
```


MSE of error variance $\sigma^2_{X,j}$, $\sigma^2_{Y,j}$:

```{r}
# X
MSE_s2 = matrix(data = NA, nrow = length(s2_x_save), ncol = J)
for (i in 1:length(s2_x_save)) {
  for (j in 1:J) {
    MSE_s2[i,j] = (s2_x[j] - s2_x_save[[i]][j])^2 / s2_x[j]^2
  }
}
print(paste0("The mean MSE for the error variances {", 
             symbol("sigma") %p% supsc("2"), "X, ", "are:"))
print(round(colMeans(MSE_s2, na.rm = TRUE), digits = 3))

# Y
MSE_s2 = matrix(data = NA, nrow = length(s2_y_save), ncol = J)
for (i in 1:length(s2_y_save)) {
  for (j in 1:J) {
    MSE_s2[i,j] = (s2_y[j] - s2_y_save[[i]][j])^2 / s2_y[j]^2
  }
}
print(paste0("The mean MSE for the error variances {", 
             symbol("sigma") %p% supsc("2"), "Y, ", "are:"))
print(round(colMeans(MSE_s2, na.rm = TRUE), digits = 3))
```


MSE of outcome-specific scaling parameters $\beta^x$, $\beta^y$

```{r}
# X
beta_MSE = matrix(data = NA, nrow = length(beta.X_save), ncol = J)
for (i in 1:length(beta.X_save)) {
  for (j in 1:J) {
    beta_MSE[i,] = (beta.X - beta.X_save[[i]])^2 / (beta.X)^2
  }
}

print("Mean MSE for the outcome-specific scaling parameters in X are:")
print(paste(symbol("beta") %p% supsc("x"), ":", 
            round(colMeans(as.matrix(beta_MSE, na.rm = TRUE)), digits = 3)))

# Y
beta_MSE = matrix(data = NA, nrow = length(beta.Y_save), ncol = J)
for (i in 1:length(beta.Y_save)) {
  for (j in 1:J) {
    beta_MSE[i,] = (beta.Y - beta.Y_save[[i]])^2 / (beta.Y)^2
  }
}

print("Mean MSE for the outcome-specific scaling parameters in Y are:")
print(paste(symbol("beta") %p% supsc("y"), ":", 
            round(colMeans(as.matrix(beta_MSE, na.rm = TRUE)), digits = 3)))
```


Calculate MSDE for latent component estimates and MSE for score variance estimates, and plot true and 
estimated trajectories with (5th, 50th, 95th) MSDE values.

*Subject-level*

$\psi_\ell(t)$
```{r}
MSDE_Psi = matrix(data = NA, nrow = length(Psi_save), ncol = L)
Psi_1_est = matrix(data = NA, nrow = length(Psi_save), ncol = N)
Omega_x_est = matrix(data = NA, nrow = length(Psi_save), ncol = L)
Omega_y_est = matrix(data = NA, nrow = length(Psi_save), ncol = L)
for (i in 1:length(Psi_save)) {
  Psi_est = Psi_save[[i]]
  if(NCOL(Psi_est) >= L){ # skipped if estimated shared components are less than the true ones.
    if(NCOL(Psi_est) == 1){ # L = L_est = 1, the correspondence is fixed.
      MSDE_Psi[i,1] = min(
        trapz(tobs, (Psi - Psi_est)^2)/trapz(tobs, Psi^2), 
        trapz(tobs, (Psi + Psi_est)^2)/trapz(tobs, Psi^2))
      Psi_1_est[i,] = Psi_est
      Omega_x_est[i,1] = Omega_x_save[[i]]
      Omega_y_est[i,1] = Omega_y_save[[i]]
    }else{ # L_est > L = 1
      # When surplus shared component estimates are returned, the one having the 
      # lowest MSDE with the true component is referred to as its correspondence.
      for (j in 1:NCOL(Psi_est)) {
        MSDE = min(trapz(tobs, (Psi - Psi_est[,j])^2) / 
                     trapz(tobs, Psi^2), 
                 trapz(tobs, (Psi + Psi_est[,j])^2) / 
                   trapz(tobs, Psi^2))
        if(j > 1){
          if(MSDE < MSDE_Psi[i, 1]){
            MSDE_Psi[i, 1] = MSDE
            Psi_1_est[i,] = Psi_est[,j]
            Omega_x_est[i,1] = Omega_x_save[[i]][j,j]
            Omega_y_est[i,1] = Omega_y_save[[i]][j,j]
          }
        }else{
          MSDE_Psi[i, 1] = MSDE
          Psi_1_est[i,] = Psi_est[,j]
          Omega_x_est[i,1] = Omega_x_save[[i]][j,j]
          Omega_y_est[i,1] = Omega_y_save[[i]][j,j]
        }
      }
    }
  }
}

print(paste0("The mean MSDE for the shared components {", symbol("psi"), 
             "(t)} ", "are:"))
print(round(colMeans(MSDE_Psi, na.rm = TRUE), digits = 3))


# Plots
## Choose the estimate with (5th, 50th, 95th) MSDE with the true latent component
quantile_mat = matrix(nrow = L, ncol = 3) # a matrix saves the (5th, 50th, 95th) MSDE values.
for(j.true in 1:L){
  quantile_mat[j.true,] = sort(na.omit(MSDE_Psi)[,j.true])[
    round(nrow(na.omit(MSDE_Psi)) * c(.05, .50, .95))] 
}

ind_mat = matrix(nrow = L, ncol = 3) # a matrix contains the indices corresponding to the selected MSDE
for(j.true in 1:L){
  for (j in 1:3) {
    ind_mat[j.true,j] = which.min(abs(MSDE_Psi[,j.true] - quantile_mat[j.true,j]))
  }
}

## plot for psi_1(t)
psi_1_5th = Psi_1_est[ind_mat[1,1],]
psi_1_50th = Psi_1_est[ind_mat[1,2],]
psi_1_95th = Psi_1_est[ind_mat[1,3],]

par(mar = c(5.1, 5, 4.1, 2.1))
matplot(tobs, cbind(Psi[,1], psi_1_5th, psi_1_50th, psi_1_95th), 
        type = "l", lwd = 3, xlab = "Time", ylab = "",  cex.lab = 1.9, 
        cex.axis = 1.5, col=c("black", "red","black", "blue"), lty = c(1,2,3,4), 
        mgp = c(2.5,1,0), xaxs = "i") #, ylim = c(-.25, 1.75)
title(main = list("", cex = 2.5), adj = 0, line = 1.5)
mtext(expression(hat(psi)[1]*(t)), side=2, line=1.9, cex=2)
legend("bottomleft", c("true", "5th", "50th", "95th"), lty = c(1,2,3,4), 
       col=c("black", "red", "black", "blue"), lwd = 2, cex=1.2)
```


$\phi_k(t)$
```{r}
MSDE_Phi = matrix(data = NA, nrow = length(Phi_save), ncol = K)
Phi_1_est = matrix(data = NA, nrow = length(Phi_save), ncol = N)
Theta_est = matrix(data = NA, nrow = length(Theta_save), ncol = K)
for (i in 1:length(Phi_save)) {
  Phi_est = Phi_save[[i]]
  if(NCOL(Phi_est) >= K){
    if(NCOL(Phi_est) == 1){ # K = K_est = 1.
      MSDE_Phi[i,1] = min(
        trapz(tobs, (Phi - Phi_est)^2) / trapz(tobs, Phi^2), 
        trapz(tobs, (Phi + Phi_est)^2) / trapz(tobs, Phi^2))
      Phi_1_est[i,] = Phi_est
      Theta_est[i,1] = Theta_save[[i]]
    }else{ # K_est > K = 1
      for (j in 1:NCOL(Phi_est)) {
        MSDE = min(trapz(tobs, (Phi - Phi_est[,j])^2) / 
                     trapz(tobs, Phi^2), 
                 trapz(tobs, (Phi + Phi_est[,j])^2) / 
                   trapz(tobs, Phi^2))
        if(j > 1){
          if(MSDE < MSDE_Phi[i, 1]){
            MSDE_Phi[i, 1] = MSDE
            Phi_1_est[i,] = Phi_est[,j]
            Theta_est[i,1] = Theta_save[[i]][j,j]
          }
        }else{
          MSDE_Phi[i, 1] = MSDE
          Phi_1_est[i,] = Phi_est[,j]
          Theta_est[i,1] = Theta_save[[i]][j,j]
        }
      }
    }
  }
}

print(paste0("The mean MSDE for the unique components {", symbol("phi"), 
             "(t)} ", "are:"))
print(round(colMeans(MSDE_Phi, na.rm = TRUE), digits = 3))


# Plots
quantile_mat = matrix(nrow = K, ncol = 3)
for(j.true in 1:K){
  quantile_mat[j.true,] = sort(na.omit(MSDE_Phi)[,j.true])[
    round(nrow(na.omit(MSDE_Phi)) * c(.05, .50, .95))] 
}
ind_mat = matrix(nrow = K, ncol = 3) 
for(j.true in 1:K){
  for (j in 1:3) {
    ind_mat[j.true,j] = which.min(abs(MSDE_Phi[,j.true] - quantile_mat[j.true,j]))
  }
}

## plot for phi_1(t)
phi_1_5th = Phi_1_est[ind_mat[1,1],]
phi_1_50th = Phi_1_est[ind_mat[1,2],]
phi_1_95th = Phi_1_est[ind_mat[1,3],]

par(mar = c(5.1, 5, 4.1, 2.1))
matplot(tobs, cbind(Phi[,1], phi_1_5th, phi_1_50th, phi_1_95th), 
        type = "l", lwd = 3, xlab = "Time", ylab = "", cex.lab = 1.9, 
        cex.axis = 1.5, col=c("black", "red","black", "blue"), lty = c(1,2,3,4), 
        mgp = c(2.5,1,0), , xaxs = "i")
title(main = list("", cex = 2.5), adj = 0, line = 1.5)
mtext(expression(hat(phi)[1]*(t)), side=2, line=1.9, cex=2)
```


$\gamma_r(t)$
```{r}
MSDE_Gamma = matrix(data = NA, nrow = length(Gamma_save), ncol = R)
Gamma_1_est = matrix(data = NA, nrow = length(Gamma_save), ncol = N)
Lambda_est = matrix(data = NA, nrow = length(Lambda_save), ncol = R)
for (i in 1:length(Gamma_save)) {
  Gamma_est = Gamma_save[[i]]
  if(NCOL(Gamma_est) >= R){ 
    if(NCOL(Gamma_est) == 1){ # R = R_est = 1.
      MSDE_Gamma[i,1] = min(
        trapz(tobs, (Gamma - Gamma_est)^2) / trapz(tobs, Gamma^2), 
        trapz(tobs, (Gamma + Gamma_est)^2) / trapz(tobs, Gamma^2))
      Gamma_1_est[i,] = Gamma_est
      Lambda_est[i,1] = Lambda_save[[i]]
    }else{ # R_est > R = 1
      for (j in 1:NCOL(Gamma_est)) {
        MSDE = min(trapz(tobs, (Gamma - Gamma_est[,j])^2) / 
                     trapz(tobs, Gamma^2), 
                 trapz(tobs, (Gamma + Gamma_est[,j])^2) / 
                   trapz(tobs, Gamma^2))
        if(j > 1){
          if(MSDE < MSDE_Gamma[i, 1]){
            MSDE_Gamma[i,1] = MSDE
            Gamma_1_est[i,] = Gamma_est[,j]
            Lambda_est[i,1] = Lambda_save[[i]][j,j]
          }
        }else{
          MSDE_Gamma[i,1] = MSDE
          Gamma_1_est[i,] = Gamma_est[,j]
          Lambda_est[i,1] = Lambda_save[[i]][j,j]
        }
      }
    }
  }
}

print(paste0("The mean MSDE for the unique components {", symbol("gamma"), 
             "(t)} ", "are:"))
print(round(colMeans(MSDE_Gamma, na.rm = TRUE), digits = 3))


# Plots
quantile_mat = matrix(nrow = R, ncol = 3)
for(j.true in 1:R){
  quantile_mat[j.true,] = sort(na.omit(MSDE_Gamma)[,j.true])[
    round(nrow(na.omit(MSDE_Gamma)) * c(.05, .50, .95))] 
}

ind_mat = matrix(nrow = R, ncol = 3)
for(j.true in 1:R){
  for (j in 1:3) {
    ind_mat[j.true,j] = which.min(abs(MSDE_Gamma[,j.true] - quantile_mat[j.true,j]))
  }
}

## plot for gamma_1(t)
gamma_1_5th = Gamma_1_est[ind_mat[1,1],]
gamma_1_50th = Gamma_1_est[ind_mat[1,2],]
gamma_1_95th = Gamma_1_est[ind_mat[1,3],]

par(mar = c(5.1, 5, 4.1, 2.1))
matplot(tobs, cbind(Gamma[,1], gamma_1_5th, gamma_1_50th, gamma_1_95th), 
        type = "l", lwd = 3, xlab = "Time", ylab = "", cex.lab = 1.9, 
        cex.axis = 1.5, col=c("black", "red","black", "blue"), lty = c(1,2,3,4),
        mgp = c(2.5,1,0), xaxs = "i")
title(main = list("", cex = 2.5), adj = 0, line = 1.5)
mtext(expression(hat(gamma)[1]*(t)), side=2, line=1.9, cex=2)
```


Calculate MSE of score variances

```{r}
scoreVar_MSE = matrix(data = NA, nrow = length(Omega_x_save), ncol = (2*L + K + R))
for (i in 1:length(Omega_x_save)) {
    scoreVar_MSE[i,1:L] = (diag(Omega_x) - Omega_x_est[i,])^2 / (diag(Omega_x))^2
    scoreVar_MSE[i,(L+1):(L+L)] = (diag(Omega_y) - Omega_y_est[i,])^2 / (diag(Omega_y))^2
    scoreVar_MSE[i,(2*L+1):(2*L+K)] = (diag(Theta) - Theta_est[i,])^2 / (diag(Theta))^2
    scoreVar_MSE[i,(2*L+K+1):(2*L+K+R)] = (diag(Lambda) - Lambda_est[i,])^2 / (diag(Lambda))^2
}

print("Mean MSE for the following score variances are:")
print(paste(symbol("omega") %p% supsc("x"), ":", 
            round(colMeans(as.matrix(scoreVar_MSE[,1:L]), na.rm = TRUE), digits = 3)))
print(paste(symbol("omega") %p% supsc("y"), ":", 
            round(colMeans(as.matrix(scoreVar_MSE[,(L+1):(L+L)]), na.rm = TRUE), digits = 3)))
print(paste(symbol("theta"), ":", 
            round(colMeans(as.matrix(scoreVar_MSE[,(2*L+1):(2*L+K)]), na.rm = TRUE), digits = 3)))
print(paste(symbol("lambda"), ":", 
            round(colMeans(as.matrix(scoreVar_MSE[,(2*L+K+1):(2*L+K+R)]), na.rm = TRUE), digits = 3)))
```


*Outcome-level*

$\tau^x_v(t)$
```{r}
MSDE_tau_x = matrix(data = NA, nrow = length(tau_x_save), ncol = V_x)
tau_x_1_est = matrix(data = NA, nrow = length(tau_x_save), ncol = N)
tau_x_2_est = matrix(data = NA, nrow = length(tau_x_save), ncol = N)
varsigma_x_est = matrix(data = NA, nrow = length(varsigma_x_save), ncol = V_x)
for (i in 1:length(tau_x_save)) {
  tau_x_est = tau_x_save[[i]]
  if(NCOL(tau_x_est) >= V_x){
    min.ind = c()
    for(j.true in 1:V_x){
      MSDE = c()
      for (j in 1:NCOL(tau_x_est)) {
        MSDE = append(MSDE, min(
          trapz(tobs, (tau_x[,j.true] - tau_x_est[,j])^2) / 
            trapz(tobs, tau_x[,j.true]^2),
          trapz(tobs, (tau_x[,j.true] + tau_x_est[,j])^2) / 
            trapz(tobs, tau_x[,j.true]^2)))
      }
      min.ind = append(min.ind, which.min(MSDE))  
      MSDE_tau_x[i, j.true] = min(MSDE)
      varsigma_x_est[i,j.true] = varsigma_x_save[[i]][which.min(MSDE)]
      if(j.true == 1){tau_x_1_est[i,] = tau_x_est[,which.min(MSDE)]}
      if(j.true == 2){tau_x_2_est[i,] = tau_x_est[,which.min(MSDE)]}
    }
  }
}

print(paste0("The mean MSDE for the unique components {", symbol("tau_x"), 
             "(t)} ", "are:"))
print(round(colMeans(MSDE_tau_x, na.rm = TRUE), digits = 3))


# Plots
quantile_mat = matrix(nrow = V_x, ncol = 3)
for(j.true in 1:V_x){
  quantile_mat[j.true,] = sort(na.omit(MSDE_tau_x)[,j.true])[
    round(nrow(na.omit(MSDE_tau_x)) * c(.05, .50, .95))] 
}
ind_mat = matrix(nrow = V_x, ncol = 3) 
for(j.true in 1:V_x){
  for (j in 1:3) {
    ind_mat[j.true,j] = which.min(abs(MSDE_tau_x[,j.true] - quantile_mat[j.true,j]))
  }
}

## plot for tau^x_1(t)
tau_x_1_5th = tau_x_1_est[ind_mat[1,1],]
tau_x_1_50th = tau_x_1_est[ind_mat[1,2],]
tau_x_1_95th = tau_x_1_est[ind_mat[1,3],]

par(mar = c(5.1, 5, 4.1, 2.1))
matplot(tobs, cbind(tau_x[,1], tau_x_1_5th, tau_x_1_50th, tau_x_1_95th), 
        type = "l", lwd = 3, xlab = "Time", ylab = "", cex.lab = 1.9, 
        cex.axis = 1.5, col=c("black", "red","black", "blue"), lty = c(1,2,3,4), 
        mgp = c(2.5,1,0), , xaxs = "i", ylim = c(-.25, 1.75))
title(main = list("", cex = 2.5), adj = 0, line = 1.5)
mtext(expression(hat(tau)[1]^{x}*(t)), side=2, line=1.9, cex=2)



## plot for tau^x_2(t)
tau_x_2_5th = tau_x_2_est[ind_mat[2,1],]
tau_x_2_50th = tau_x_2_est[ind_mat[2,2],]
tau_x_2_95th = tau_x_2_est[ind_mat[2,3],]

par(mar = c(5.1, 5, 4.1, 2.1))
matplot(tobs, cbind(tau_x[,2], tau_x_2_5th, tau_x_2_50th, tau_x_2_95th), 
        type = "l", lwd = 3, xlab = "Time", ylab = "",  cex.lab = 1.9, 
        cex.axis = 1.5, col=c("black", "red","black", "blue"), 
        lty = c(1,2,3,4), mgp = c(2.5,1,0), , xaxs = "i")
title(main = list("", cex = 2.5), adj = 0, line = 1.5)
mtext(expression(hat(tau)[2]^{x}*(t)), side=2, line=1.9, cex=2)
```


$\tau^y_v(t)$
```{r}
MSDE_tau_y = matrix(data = NA, nrow = length(tau_y_save), ncol = V_y)
tau_y_1_est = matrix(data = NA, nrow = length(tau_y_save), ncol = N)
tau_y_2_est = matrix(data = NA, nrow = length(tau_y_save), ncol = N)
tau_y_3_est = matrix(data = NA, nrow = length(tau_y_save), ncol = N)
varsigma_y_est = matrix(data = NA, nrow = length(varsigma_y_save), ncol = V_y)
for (i in 1:length(tau_y_save)) {
  tau_y_est = tau_y_save[[i]]
  if(NCOL(tau_y_est) >= V_y){
    min.ind = c()
    for(j.true in 1:V_y){
      MSDE = c()
      for (j in 1:NCOL(tau_y_est)) {
        MSDE = append(MSDE, min(
          trapz(tobs, (tau_y[,j.true] - tau_y_est[,j])^2) / 
            trapz(tobs, tau_y[,j.true]^2),
          trapz(tobs, (tau_y[,j.true] + tau_y_est[,j])^2) / 
            trapz(tobs, tau_y[,j.true]^2)))
      }
      min.ind = append(min.ind, which.min(MSDE))  
      MSDE_tau_y[i, j.true] = min(MSDE)
      varsigma_y_est[i,j.true] = varsigma_y_save[[i]][which.min(MSDE)]
      if(j.true == 1){tau_y_1_est[i,] = tau_y_est[,which.min(MSDE)]}
      if(j.true == 2){tau_y_2_est[i,] = tau_y_est[,which.min(MSDE)]}
      if(j.true == 3){tau_y_3_est[i,] = tau_y_est[,which.min(MSDE)]}
    }
  }
}

print(paste0("The mean MSDE for the unique components {", symbol("tau_y"), 
             "(t)} ", "are:"))
print(round(colMeans(MSDE_tau_y, na.rm = TRUE), digits = 3))


# Plots
quantile_mat = matrix(nrow = V_y, ncol = 3)
for(j.true in 1:V_y){
  quantile_mat[j.true,] = sort(na.omit(MSDE_tau_y)[,j.true])[
    round(nrow(na.omit(MSDE_tau_y)) * c(.05, .50, .95))] 
}
ind_mat = matrix(nrow = V_y, ncol = 3)
for(j.true in 1:V_y){
  for (j in 1:3) {
    ind_mat[j.true,j] = which.min(abs(MSDE_tau_y[,j.true] - quantile_mat[j.true,j]))
  }
}

## plot for tau^y_1(t)
tau_y_1_5th = tau_y_1_est[ind_mat[1,1],]
tau_y_1_50th = tau_y_1_est[ind_mat[1,2],]
tau_y_1_95th = tau_y_1_est[ind_mat[1,3],]

par(mar = c(5.1, 5, 4.1, 2.1))
matplot(tobs, cbind(tau_y[,1], tau_y_1_5th, tau_y_1_50th, tau_y_1_95th), 
        type = "l", lwd = 3, xlab = "Time", ylab = "", cex.lab = 1.9, 
        cex.axis = 1.5, col=c("black", "red","black", "blue"), lty = c(1,2,3,4), 
        mgp = c(2.5,1,0), , xaxs = "i", ylim = c(-.25, 1.75))
title(main = list("", cex = 2.5), adj = 0, line = 1.5)
mtext(expression(hat(tau)[1]^{y}*(t)), side=2, line=1.9, cex=2)

## plot for tau^y_2(t)
tau_y_2_5th = tau_y_2_est[ind_mat[2,1],]
tau_y_2_50th = tau_y_2_est[ind_mat[2,2],]
tau_y_2_95th = tau_y_2_est[ind_mat[2,3],]

par(mar = c(5.1, 5, 4.1, 2.1))
matplot(tobs, cbind(tau_y[,2], tau_y_2_5th, tau_y_2_50th, tau_y_2_95th), 
        type = "l", lwd = 3, xlab = "Time", ylab = "",  cex.lab = 1.9, 
        cex.axis = 1.5, col=c("black", "red","black", "blue"), 
        lty = c(1,2,3,4), mgp = c(2.5,1,0), , xaxs = "i")
title(main = list("", cex = 2.5), adj = 0, line = 1.5)
mtext(expression(hat(tau)[2]^{y}*(t)), side=2, line=1.9, cex=2)

## plot for tau_y_3(t)
tau_y_3_5th = tau_y_3_est[ind_mat[3,1],]
tau_y_3_50th = tau_y_3_est[ind_mat[3,2],]
tau_y_3_95th = tau_y_3_est[ind_mat[3,3],]

par(mar = c(5.1, 5, 4.1, 2.1))
matplot(tobs, cbind(tau_y[,3], tau_y_3_5th, tau_y_3_50th, tau_y_3_95th), 
        type = "l", lwd = 3, xlab = "Time", ylab = "",  cex.lab = 1.9, 
        cex.axis = 1.5, col=c("black", "red","black", "blue"), 
        lty = c(1,2,3,4), mgp = c(2.5,1,0), , xaxs = "i")
title(main = list("", cex = 2.5), adj = 0, line = 1.5)
mtext(expression(hat(tau)[3]^{y}*(t)), side=2, line=1.9, cex=2)
```


Calculate MSE of score variances


```{r}
scoreVar_MSE = matrix(data = NA, nrow = length(varsigma_x_save), ncol = (V_x + V_y))
for (i in 1:length(varsigma_x_save)) {
    scoreVar_MSE[i,1:V_x] = (diag(varsigma_x) - varsigma_x_est[i,])^2 / (diag(varsigma_x))^2
    scoreVar_MSE[i,(V_x+1):(V_x+V_y)] = (diag(varsigma_y) - varsigma_y_est[i,])^2 / (diag(varsigma_y))^2
}

print("Mean MSE for the following score variances are:")
print(paste(symbol("varsigma") %p% supsc("x"), ":", 
            round(colMeans(as.matrix(scoreVar_MSE[,1:V_x]), na.rm = TRUE), digits = 3)))
print(paste(symbol("varsigma") %p% supsc("y"), ":", 
            round(colMeans(as.matrix(scoreVar_MSE[,(V_x+1):(V_x+V_y)]), na.rm = TRUE),digits = 3)))
```










